Benchmark!!1

### 2021-01-14 paper/main.tex
style improvement
more description of implementation-specific quirks

\section{Languages and implementations considered}
\subsection{C++}
\subsection{Python}
\subsection{Java}
\subsection{Julia}
\subsection{Rust}
\subsection{JavaScript}
JavaScript is not expected to perform as well as, say, C++. However
there are several reasons why it can be appealing to implement routing
algorithms in JavaScript:
\begin{itemize}
\item Any program written in JavaScript can run on virtually any
  computer with a modern web browser, including smartphones.
\item Integrating JavaScript in a web interface is especially easy.
\item JavaScript engines have received considerable attention from
  major companies and are the subject of fierce competition on
  performance. As a result they have seen vast performance
  improvements over the years and the trend is likely to continue.
\end{itemize}

@@

@@

\subsection{Something DP-like}

@@

\section{Language-specific notes}
\subsection{Python}
Python does not allow to inline functions. Since using a flat matrix
requires to wrap matrix value lookup in a function, using a flat distance
matrix in Python is not beneficial. It would of course be possible to
inline by hand, i.e. formulate the correct 1-dimensional index in the
matrix for every lookup, but that would be impractical and defeat the
purposes and philosophy stated in Section~\ref{sec:intro}.
Running Python code that uses a flat representation with a JIT compiler
(e.g. Pypy, Numba) introduces automatic inlining and is likely to
remedy this.
\subsection{Julia}
Julia has native multi-dimensional arrays, which may be more efficient
than a flat representation. We will determine which version to use
based on experiment. Since Julia's arrays start with index 1, the code
for flat representation lookup of the distance between vertices $i$
and $j$ in matrix $d$ is \lstinline{d[i*n-n+j]}. It is worth noting
that it involves one subtraction on top of the addition and
multiplication used for 0-indexed languages.
\subsection{JavaScript}
All benchmarks are repetitive in nature, as they successively perform similar
operations with different input data. In JavaScript, when our
implementation loads 40 instances and performs the same
benchmarking operations on all 40 them, we observe a significant performance
hit after a few instances (2-5 times slower). In order to remedy this,
we re-start the program individually for each instance. Advanced
knowledge of this specific JavaScript engine might allow to remove that
performance hit, but this is clearly outside the scope of this
study. However it is worth noting that this can be a concern in general.
\subsection{Numba}
Numba is not yet feature-complete for Numpy. A direct consequence is that some
benchmarks cannot be implemented for Numba. For example $np.insert()$ is not
implemented, but this function is needed for the $LN$
benchmark. Therefore the $LNS$ benchmark cannot be implemented with
Numba. Still, Numba is significant enough to be benchmarked and its
current performance are an indication of what to expect in years to come.

@@

>>>>

\section{Languages and implementations considered}
\subsection{C++}
Two implementations: c++14 vs c++98.

\subsection{Python}
Python vs Pypy vs Numpy vs Numba.

\subsection{Java}

\subsection{Julia}
General performance tips: read the specific doc section! One unique
official document outlines how to write efficient code.

\subsection{Rust}

\subsection{JavaScript}
JavaScript is not expected to perform as well as, say, C++. However
there are several reasons why it can be appealing to implement routing
algorithms in JavaScript:
\begin{itemize}
\item Any program written in JavaScript can run on virtually any
  computer with a modern web browser, including smartphones.
\item Integrating JavaScript in a web interface is especially easy.
\item JavaScript engines have received considerable attention from
  major companies and are the subject of fierce competition on
  performance. As a result they have seen vast performance
  improvements over the years and the trend is likely to continue.
\end{itemize}

@@

\subsection{General cross-language considerations}
Subjective considerations:
\begin{itemize}
\item Typically easier to program using Python, Julia or JavaScript.
\item Typically harder to program using Rust than anything else but
  might be due to lack of experience. However my experience in
  JavaScript is much lower than in C++ and I still did everything
  faster in JavaScript than in C++.
\end{itemize}

@@

\subsection{Dynamic programming for ESPPRC}
Motivation: column generation.

Mention here espprc-index and how rust does not allow multiple
references to other labels.

\subsection{Maximum flow problem}
Motivation: branch-and-cut.

@@

\section{Implementation-specific notes}
\subsection{Python}
Python does not allow to inline functions. Since using a flat matrix
requires to wrap matrix value lookup in a function, using a flat distance
matrix in Python is not beneficial. It would of course be possible to
inline by hand, i.e. formulate the correct 1-dimensional index in the
matrix for every lookup, but that would be impractical and defeat the
purposes and philosophy stated in Section~\ref{sec:intro}.
Running Python code that uses a flat representation with a JIT compiler
(e.g. Pypy, Numba) introduces automatic inlining and is likely to
remedy this.
\subsection{Julia}
Julia has native multi-dimensional arrays, which may be more efficient
than a flat representation. We will determine which version to use
based on experiment. Since Julia's arrays start with index 1, the code
for flat representation lookup of the distance between vertices $i$
and $j$ in matrix $d$ is \lstinline{d[i*n-n+j]}. It is worth noting
that it involves one subtraction on top of the addition and
multiplication used for 0-indexed languages.

Using two threads instead of 1 improves performance by about 10\%,
although it is unclear what the second thread is used for.

\subsection{JavaScript}
All benchmarks are repetitive in nature, as they successively perform similar
operations with different input data. In JavaScript, when our
implementation loads 40 instances and performs the same
benchmarking operations on all 40 them, we observe a significant performance
hit after a few instances (2-5 times slower). In order to remedy this,
we re-start the program individually for each instance. Advanced
knowledge of this specific JavaScript engine might allow to remove that
performance hit, but this is clearly outside the scope of this
study. However it is worth noting that this can be a concern in general.

\subsection{Numba}
Numba is not yet feature-complete for Numpy. A direct consequence is that some
benchmarks cannot be implemented for Numba. For example
\lstinline{np.insert()} is not implemented, but this function is
needed for the $LN$ benchmark. Therefore the $LNS$ benchmark cannot be
implemented with Numba. Still, Numba is significant enough to be
benchmarked and its current performance are an indication of what to
expect in years to come.

\subsection{Java}
Mention inconsistencies between Java versions: code written for Java 8
does not compile with Java 11. Because Pair is not in the API any more.

\subsection{Rust}
Issues:
\begin{itemize}
\item Changing from i32 to i64 is problematic somehow
\item Not possible to have refs to both successors and predecessor in
  Label, hence had to handle indices by hand in specific espprc implementation
\end{itemize}

@@

\section*{Acknowledgements}
Sebastian for advice and help with Rust.
